---
title: "Calibrating Fall Run DSM"
output: html_notebook
---

# TODO change this to be a markdown 
# look at the habitat to see how the yaml and vignettes work for the docs website

## Objective

We are using genetic algorithms, a type of stochastic optimization, to calibrate the following model coefficients:

* `..surv_adult_enroute_int` used in submodel `surv_adult_enroute`
* `..surv_adult_prespawn_int` used in submodel `surv_adult_prespawn`
* `..surv_egg_to_fry_int` used in submodel `surv_egg_to_fry`
* `..surv_juv_rear_int` used in submodel `surv_juv_rear`
* `..surv_juv_rear_contact_points` used in submodel `surv_juv_rear`
* `..surv_juv_rear_prop_diversions` used in submodel `surv_juv_rear`
* `..surv_juv_rear_total_diversions` used in submodel `surv_juv_rear`
* `..surv_juv_bypass_int` used in submodel `surv_juv_bypass`
* `..surv_juv_delta_int` used in submodel `surv_juv_delta`
* `..surv_juv_delta_contact_points` used in submodel `surv_juv_delta`
* `..surv_juv_delta_total_diverted` used in submodel `surv_juv_delta`
* `..surv_juv_outmigration_sj_int` used in submodel `surv_juv_outmigration_san_joaquin`
* `..ocean_entry_success_int`  used in submodel `ocean_entry_success`

# TODO add link to the documentaiton website

We are restricting the calibration to watersheds with Grandtab data, these are:

```{r, echo=FALSE}
keep <- c(1,6,7,10,12,19,20,23,26:30)
fallRunDSM::watershed_labels[keep]
```


## Packages

```{r, message=FALSE}
library(GA)
library(fallRunDSM)
library(DSMCalibrationData)
library(tidyverse)
```

We are using the [GA](http://luca-scr.github.io/GA/index.html) R package to implement our genetic algorithms search to calibrate the above intercepts for the `fallRunDSM::fall_run_model()` function of the [fallRunDSM](https://cvpia-osc.github.io/fallRunDSM) R package.

Additionally, we are using [DSMCalibrationData](https://github.com/CVPIA-OSC/DSMCalibrationData) to access calibration related data.

## Calibrating Data

### GrandTab

We will be minimizing the error between [GrandTab](https://wildlife.ca.gov/Conservation/Fishes/Chinook-Salmon/Anadromous-Assessment) estimated escapement data for the years 1998-2017 to natural spawners predicted by the DSM. We are further scaling this error by a weight representing the amount of observed data available for a given watershed. Given that
GrandTab includes both hatchery and natural spawners, we will scale GrandTab escapement
values using the predicted proportion natural spawners for a given year generated by the model.
This will allow us to compare predicted natural spawners to psuedo GrandTab natural spawners.

For calibration, we have prepared the GrandTab data in two ways:

1. `DSMCalibrationData::grandtab_observed` - This data will be used in the fitness function to measure the difference between model predictions and observed escapement. Missing values are NA and we made all records less than 100 NA to account for a lack of confidence for counts less than 100. 

2. `DSMCalibrationData::grandtab_imputed` - This data will be used to calculate the number of juveniles during the 20 year simulation. The GrandTab data is incomplete for many watersheds during the 20 year period of calibration. For watersheds with no GrandTab data, we used 40 as the default escapement value. For watersheds with incomplete data, we used the mean escapement value. 

# TODO change these to plots instead of tables

```{r}
grandtab_observed$fall
```

```{r}
grandtab_imputed$fall
```

### Selected Proxy Years

The data inputs to the DSM are for years 1980-1999. We selected proxy years for 1998-2017
from the 1980-1999 model inputs by [comparing the DWR water year indices](https://cdec.water.ca.gov/reportapp/javareports?name=WSIHIST) in this [analysis](https://github.com/CVPIA-OSC/DSMCalibrationData/blob/main/data-raw/synthetic_years.R).

These inputs are generated by calling `DSMCalibrationData::set_synth_years(params)`, with a `params` object from the life cycle model ex: `fallRunDSM::params`. See the documentation for more details on the [DSMCalibrationData](https://github.com/CVPIA-OSC/DSMCalibrationData)
R package.

#### Spawning Inputs Proxy Years
```{r echo=FALSE}
tibble::tibble(
  `Calibration Year` = names(DSMCalibrationData::calibration_year_spawn_index), 
  `Selected Proxy Year` = DSMCalibrationData::calibration_year_spawn_index) %>% 
  knitr::kable()
```

#### Rearing Inputs Proxy Years
```{r echo=FALSE}
tibble::tibble(
  `Calibration Year` = names(DSMCalibrationData::calibration_year_index), 
  `Selected Proxy Year` = DSMCalibrationData::calibration_year_index) %>% 
  knitr::kable()
```


## Fitness Function

We are minimizing the sum of squared errors of simulated natural adults and grandtab 
adults scaled using simulated proportion natural, normalized by mean escapement. The
purpose of normalizing is to prevent watersheds with large escapement values from
dominating the calibration. In addition to normalizing we are adding weights to 
each error term corresponding to the amount data available at that watershed. Doing
so will penalize watersheds with more data whose error term is large, but be more 
lineant with watersheds with less data.

All intercepts and watersheds that need calibration are described in the Objective
section of this document.

Note, `..surv_adult_enroute_int` and `..ocean_entry_success_int` are vectors and expanded within the fitness function arguments.

```{r}
# Fitness Function ------------------
fall_run_fitness <- function(
  known_adults,
  seeds,
  params,
  surv_adult_enroute,
  surv_adult_prespawn,
  surv_egg_to_fry,
  bypass_surv_juv,
  upsac_surv_juv,
  butte_surv_juv,
  clear_surv_juv,
  deer_surv_juv,
  mill_surv_juv,
  sac_surv_juv,
  feather_and_bear_surv_juv,
  yuba_surv_juv,
  american_surv_juv,
  deltatribs_surv_juv,
  moke_surv_juv,
  merced_surv_juv,
  stan_surv_juv,
  tuol_surv_juv,
  sj_surv_juv,
  surv_juv_rear_contact_points,
  surv_juv_rear_prop_diversions,
  surv_juv_rear_total_diversions,
  surv_juv_bypass_int,
  surv_juv_delta_int,
  surv_juv_delta_contact_points,
  surv_juv_delta_total_diverted,
  surv_juv_outmigration_sj_int,
  default_ocean_entry_surv,
  upsac_ocean_entry_surv,
  butte_ocean_entry_surv,
  deer_ocean_entry_surv,
  mill_ocean_entry_surv,
  midsactribs_ocean_entry_surv,
  yuba_ocean_entry_surv,
  american_ocean_entry_surv,
  deltatribs_ocean_entry_surv,
  moke_ocean_entry_surv,
  merced_ocean_entry_surv,
  stan_ocean_entry_surv,
  tuol_ocean_entry_surv) {

  params_init <- params

  params_init$..surv_adult_enroute_int = surv_adult_enroute
  params_init$..surv_adult_prespawn_int = surv_adult_prespawn
  params_init$..surv_egg_to_fry_int = surv_egg_to_fry
  params_init$..surv_juv_rear_int = c(`Upper Sacramento River` = upsac_surv_juv,
                                      `Antelope Creek` = deer_surv_juv,
                                      `Battle Creek` = deer_surv_juv,
                                      `Bear Creek` = deer_surv_juv,
                                      `Big Chico Creek` = deer_surv_juv,
                                      `Butte Creek` = butte_surv_juv,
                                      `Clear Creek` = clear_surv_juv,
                                      `Cottonwood Creek` = deer_surv_juv,
                                      `Cow Creek` = deer_surv_juv,
                                      `Deer Creek` = deer_surv_juv,
                                      `Elder Creek` = deer_surv_juv,
                                      `Mill Creek` = mill_surv_juv,
                                      `Paynes Creek` = deer_surv_juv,
                                      `Stony Creek` = deer_surv_juv,
                                      `Thomes Creek` = deer_surv_juv,
                                      `Upper-mid Sacramento River` = sac_surv_juv,
                                      `Sutter Bypass` = bypass_surv_juv,
                                      `Bear River` = feather_and_bear_surv_juv,
                                      `Feather River` = feather_and_bear_surv_juv,
                                      `Yuba River` = yuba_surv_juv,
                                      `Lower-mid Sacramento River` = sac_surv_juv,
                                      `Yolo Bypass` = bypass_surv_juv,
                                      `American River` = american_surv_juv,
                                      `Lower Sacramento River` = sac_surv_juv,
                                      `Calaveras River` = deltatribs_surv_juv,
                                      `Cosumnes River` = deltatribs_surv_juv,
                                      `Mokelumne River` = moke_surv_juv,
                                      `Merced River` = merced_surv_juv,
                                      `Stanislaus River` = stan_surv_juv,
                                      `Tuolumne River` = tuol_surv_juv,
                                      `San Joaquin River` = sj_surv_juv)
  params_init$..surv_juv_rear_contact_points = surv_juv_rear_contact_points
  params_init$..surv_juv_rear_prop_diversions = surv_juv_rear_prop_diversions
  params_init$..surv_juv_rear_total_diversions = surv_juv_rear_total_diversions
  params_init$..surv_juv_bypass_int = surv_juv_bypass_int
  params_init$..surv_juv_delta_int = surv_juv_delta_int
  params_init$..surv_juv_delta_contact_points = surv_juv_delta_contact_points
  params_init$..surv_juv_delta_total_diverted = surv_juv_delta_total_diverted
  params_init$..surv_juv_outmigration_sj_int = surv_juv_outmigration_sj_int
  params_init$..ocean_entry_success_int = c(
    `Upper Sacramento River` = upsac_ocean_entry_surv,
    `Antelope Creek` = default_ocean_entry_surv,
    `Battle Creek` = default_ocean_entry_surv,
    `Bear Creek` = default_ocean_entry_surv,
    `Big Chico Creek` = default_ocean_entry_surv,
    `Butte Creek` = butte_ocean_entry_surv,
    `Clear Creek` = default_ocean_entry_surv,
    `Cottonwood Creek` = default_ocean_entry_surv,
    `Cow Creek` = default_ocean_entry_surv,
    `Deer Creek` = deer_ocean_entry_surv,
    `Elder Creek` = default_ocean_entry_surv,
    `Mill Creek` = mill_ocean_entry_surv,
    `Paynes Creek` = default_ocean_entry_surv,
    `Stony Creek` = default_ocean_entry_surv,
    `Thomes Creek` = default_ocean_entry_surv,
    `Upper-mid Sacramento River` = default_ocean_entry_surv,
    `Sutter Bypass` = default_ocean_entry_surv,
    `Bear River` = midsactribs_ocean_entry_surv,
    `Feather River` = midsactribs_ocean_entry_surv,
    `Yuba River` = yuba_ocean_entry_surv,
    `Lower-mid Sacramento River` = default_ocean_entry_surv,
    `Yolo Bypass` = default_ocean_entry_surv,
    `American River` = american_ocean_entry_surv,
    `Lower Sacramento River` = default_ocean_entry_surv,
    `Calaveras River` = deltatribs_ocean_entry_surv,
    `Cosumnes River` = deltatribs_ocean_entry_surv,
    `Mokelumne River` = moke_ocean_entry_surv,
    `Merced River` = merced_ocean_entry_surv,
    `Stanislaus River` = stan_ocean_entry_surv,
    `Tuolumne River` = tuol_ocean_entry_surv,
    `San Joaquin River` = default_ocean_entry_surv)


  keep <- c(1,6,7,10,12,19,20,23,26:30)
  num_obs <- rowSums(!is.na(known_adults[keep, 6:20]))
  total_obs <- sum(!is.na(known_adults[keep, 6:20]))
  weights <- num_obs / total_obs


  tryCatch({
    preds <- fall_run_model(mode = "calibrate",
                            seeds = seeds,
                            stochastic = FALSE,
                            ..params = params_init)

    known_nats <- known_adults[keep, 6:20] * (1 - params_init$proportion_hatchery[keep])
    mean_escapent <-rowMeans(known_nats, na.rm = TRUE)

    sse <- sum(((preds[keep,] - known_nats)^2 * weights)/mean_escapent, na.rm = TRUE)

    return(sse)
  },
  error = function(e) return(1e12),
  warning = function(w) return(1e12)
  )
}
```


```{r, eval=FALSE}
library(GA)

res <- ga(type = "real-valued",
          fitness =
            function(x) -fall_run_fitness(
              known_adults = known_adults,
              seeds = calibration_seeds,
              params = params,
              x[1], x[2], x[3], x[4], x[5], x[6], x[7], x[8], x[9], x[10],
              x[11], x[12], x[13], x[14], x[15], x[16], x[17], x[18], x[19],
              x[20], x[21], x[22], x[23], x[24], x[25], x[26], x[27], x[28],
              x[29], x[30], x[31], x[32], x[33], x[34], x[35], x[36], x[37],
              x[38], x[39], x[40]
            ),
          lower = c(2.5, rep(-3.5, 39)),
          upper = rep(3.5, 40),
          popSize = 150,
          maxiter = 10000,
          run = 50,
          parallel = TRUE,
          pmutation = .4)

```


## Calibration Settings

The `ga` function allows for several paramaters to be set before calibrating a model.
Here we list the reasoning for the values passed into the `ga` functon above.

* `lower` - sets the lower bound possible for each of the parameters we are trying 
to calibrate. All parameters used in the model exist within a inverse logit function
therefore limiting the genetic algorithms space of search greatly improves its speed.
We also explicitly limit `..surv_adult_enroute` as a positive value to keep the GA 
from quickly converging to a state where all fish are dead. 

* `upper` - sets the upper bound possible for each of the parameters we are trying
to calibrate. The same reasoning as `lower`, but we add no special limitiation 
to `..surv_adult_enroute`

* `popSize` - sets population size for each iteration of the GA. We found 150 to 
be a good balance between breadth and speed of the search. 

* `maxiter` - sets the number of iterations the GA will run for. We set this value
large enough so that the `run` argument would be determining factor as to when the 
GA stopped. 

* `run` - we set this value to 50 as it allowed the algorithm to avoid local minimum
but also complete the calibration within 3-4 hours. 

* `pmutation` - set the probability of mutation, we used this value to explore
the solution space more aggresively once we had a starting position from a previous run.

The algorithm is always run in parallel. 



# Calibration Results 

add items from the run-calibration results 









