---
title: "Calibrating Fall Run DSM"
output: html_notebook
---

## Objective

We are using genetic algorithms, a type of stochastic optimization, to calibrate the following model coefficients:

* `..surv_adult_enroute_int`
* `..surv_adult_prespawn_int`
* `..surv_egg_to_fry_int`
* `..surv_juv_rear_int`
* `..surv_juv_rear_contact_points`
* `..surv_juv_rear_prop_diversions`
* `..surv_juv_rear_total_diversions`
* `..surv_juv_bypass_int`
* `..surv_juv_delta_int`
* `..surv_juv_delta_contact_points`
* `..surv_juv_delta_total_diverted`
* `..surv_juv_outmigration_sj_int`
* `..surv_juv_outmigration_sac_int_one`
* `..surv_juv_outmigration_sac_prop_diversions`
* `..surv_juv_outmigration_sac_total_diversions`
* `..surv_juv_outmigration_sac_int_two`
* `..ocean_entry_success_int`


## Packages
We are using the [GA](http://luca-scr.github.io/GA/index.html) R package to implement our genetic algorithms search
and the [fallRunDSMCalibration](https://github.com/CVPIA-OSC/fallRunDSMCalibration) R package that contains
a version of [fallRunDSM](https://github.com/CVPIA-OSC/fallRunDSM) that has exposed the above model coefficients
in the `fall_run_model` function to enable optimization of these coefficients.
Additionally, we are using [DSMCalibrationData](https://github.com/CVPIA-OSC/DSMCalibrationData)
**NEED TO BUILD DOCUMENTATION SITES FOR THESE AND UPDATE LINKS**
```{r}
library(GA)
library(fallRunDSMCalibration)
library(DSMCalibrationData)
```

## Data

```{r}
fallRunDSMCalibration::
```


## Fitness Function
```{r}

```
