---
title: "DSM Hatchery Analysis Documentation "
output: 
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{hatchery-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(tidyverse)
library(DT)
library(fallRunDSM)
```

## Description 
This analysis provides the `proportion_hatchery`, `hatchery_allocation` and `natural_adult_removal_rate` inputs to the FallRunDSM. 

### Datasource
Temporal Range: 2010 - 2012 

Most data were obtained from CWT (coded wire tag) analyses with limited data in 2010 from CVPIA ChinookProd Database 
Version 2, September 22, 2017. Data from these sources were compiled in a csv called [hatchery_fish_fate2010_2012.csv](https://dsm-docs.s3-us-west-2.amazonaws.com/adult_hatchery_documentation/hatchery_fish_fate2010_2012.csv).

**Adjustment to Feather and Yuba Rivers:**

Feather and Yuba Rivers do not separate spring and fall run
in totals. Here, we apply the proportion of spring hatchery return
adults to adjust the totals for 2010 - 2012. Data used for Feather and Yuba adjustment were from [GrandTab 2017 version.](https://dsm-docs.s3-us-west-2.amazonaws.com/adult_hatchery_documentation/GrandTab.2017.04.07.pdf)

The data table below shows the prepped data for analysis. (`No_hatch_returns` corresponds to the number of hatchery fish returning, `Hatch_flag` describes if fish are hatchery(1) or not hatchery(0)) 

```{r}
cwt_data <- read_csv(system.file("extdata", "hatchery-data", "hatchery_fish_fate2010_2012.csv", 
                                 package = "fallRunDSM", mustWork = TRUE)) %>%
  rename("Hatch_flag" = Hatch.flag)

proportion_spring_run <- dplyr::tibble(prp = c(0.076777295, 0.056932196, 0.081441457), 
                                       Year = c(2010, 2011, 2012))

# Apply fall proportions * (1- spring proportion) to feather and yuba 
fall_run_feather_yuba <- cwt_data %>%
  filter(Run == "Fall/Spr") %>%
  left_join(proportion_spring_run, by = "Year") %>%
  mutate(No_hatch_ret = No_hatch_ret * (1 - prp),
         Total_escape = Total_escape * (1 - prp),
         Run = "Fall") %>%
  select(-prp)

# Apply fall proportions * (spring proportion) to feather and yuba 
spring_run_feather_yuba <- cwt_data %>%
  filter(Run == "Fall/Spr") %>%
  left_join(proportion_spring_run, by = "Year") %>%
  mutate(No_hatch_ret = No_hatch_ret * prp,
         Total_escape = Total_escape * prp,
         Run = "Spring") %>%
  select(-prp)
  
# Remove initial "Fall/Spr" data from hatchery data and then bind new spring fall data on
cwt_data <- cwt_data %>%
  filter(Run != "Fall/Spr") %>%
  rbind(spring_run_feather_yuba, fall_run_feather_yuba)

datatable(cwt_data %>% filter(Run == "Fall") %>% select(-Run), 
          options = list(
                    searching = FALSE,
                    pageLength = 5,
                    lengthMenu = c(5, 10, 15, 20)
))
```
### Proportion Hatchery 

The `proportion_hatchery` for each tributary is calculated below. To find the `proportion_hatchery`, we grouped our data by tributary and summarized to find the mean percentage of hatchery fish for each tributary. 

```{r}
proportion_hatchery <- cwt_data %>% 
  filter(Run == "Fall") %>% 
  select(-Run) %>% 
  group_by(Trib) %>% 
  summarise(prop_hatchery = round(mean(Hatchery, na.rm = T), 3))

datatable(proportion_hatchery, 
          options = list(
                    searching = FALSE,
                    pageLength = 5,
                    lengthMenu = c(5, 10, 15, 20)
))
```



### Hatchery Allocation
The number of `hatch_adults` is determined in the `model.R` script through a multinominal distribution, with the probability equal to the `hatchery_allocation` calculated below. 

To find the `hatchery_allocation` we calculated the proportion of hatchery fish allocated to each tributary in each year, grouped our data by tributary and Run, and summarized to find the mean percentage of hatchery fish for each tributary. 

There are 5 tributaries missing data for Fall Run fish. We assign lowest proportion of Fall Run hatchery fish spawning in our observed dataset to these tributaries and we renormalize the values so that they all sum to 1.


```{r}
fish_in_wild <- cwt_data %>%
  filter(Run == "Fall") %>%
  filter(Hatch_flag == 0)

wild_totals <- fish_in_wild %>%
  group_by(Year) %>%
  summarise(total_hatch_spawn = sum(No_hatch_ret)) 

fish_in_wild <- left_join(fish_in_wild, wild_totals, by = "Year")

summarized_fish_in_wild <- fish_in_wild %>%
  mutate(trib_hatchery_percentage = No_hatch_ret/total_hatch_spawn) %>%
  group_by(Trib) %>%
  summarise(mean_trib_hatchery_percentage = mean(trib_hatchery_percentage)) %>%
  ungroup()  

hatchery_per_trib <- summarized_fish_in_wild %>%
  rbind(tibble(Trib = c("Bear Creek", "Elder Creek", "Stony Creek", "Thomes Creek", "Bear River"),
               mean_trib_hatchery_percentage = rep(min(summarized_fish_in_wild$mean_trib_hatchery_percentage), 5))) %>%
  mutate(mean_trib_hatchery_percentage = round((mean_trib_hatchery_percentage / sum(mean_trib_hatchery_percentage, 
                                                                                    na.rm = TRUE)), 4)) %>%
  rename("Fall proportion allocated to tributary" = mean_trib_hatchery_percentage)# %>%
  #mutate(across(everything(), ~replace_na(.x, 0)))

datatable(hatchery_per_trib, 
          options = list(
                    searching = FALSE,
                    pageLength = 8,
                    lengthMenu = c(8, 10, 15, 20)
)) %>% formatStyle(columns = c("Trib"), textAlign = 'center')
```

### Natural Adult Removal Rate
`natural_adult_removal_rate` is used in the fallRunDSM by the `get_spawning_adults()` function.  The number of `natural_adults_by_month` is determined by the binomial distribution with the probability equal to the `1 - natural_adult_removal_rate`. 

We calculated the natural adult removal rate for each tributary by estimating the proportion of naturally produced fish used in a hatchery (i.e., they did not spawn). 

For Fall Run `natural_adult_removal_rate`, there are 5 tributaries where fish are removed for use in hatcheries. 

```{r}
# Fall natural adult removal -------------------------------------------------------------------------------------
fall_hatchery_fish <- cwt_data %>%
  filter(Hatch_flag == 1 & Run == "Fall") %>%
  mutate(remove = Natural * Total_escape)

pairs <- matrix(c("Coleman National Fish Hatchery", "Battle Creek",
                  "Feather River Hatchery", "Feather River", "Merced River Fish Facility", 
                  "Merced River", "Nimbus Fish Hatchery", "American River", "Mokelumne Hatchery", 
                  "Mokelumne River"), 
                ncol=2, 
                byrow=T)

fall_fish_removal = NULL

fall_fish_removal <- map_df(1:nrow(pairs), function(i){
  keep_in_hatchery <- fall_hatchery_fish %>%
    filter(Trib == pairs[i, 1]) %>%
    arrange(Year)
  if (nrow(keep_in_hatchery) == 0) return()
  
  remove_from_trib <- cwt_data %>%
    filter(Run == "Fall" & Trib == pairs[i, 2]) %>%
    arrange(Year) %>%
    mutate(natural_removed = Natural * Total_escape) 
    fall_proportion_removed <- keep_in_hatchery$remove/(keep_in_hatchery$remove + remove_from_trib$natural_removed)
    fall_fish_removal <- rbind(fall_fish_removal, data.frame(Trib = rep(pairs[i, 2], 3), fall_proportion_removed, 3))


})
  
# Combine run removal rates 
natural_adult_removal_rate <- fall_fish_removal %>%
  group_by(Trib) %>%
  summarise("Fall Proportion Removed" = round(mean(fall_proportion_removed), 3))# %>%
  #mutate(across(everything(), ~replace_na(.x, 0)))

# Display it all 
datatable(natural_adult_removal_rate, 
          options = list(
                    searching = FALSE,
                    pageLength = 6,
                    lengthMenu = c(6, 10, 15, 20)
)) %>% formatStyle(columns = c("Trib"), textAlign = 'center')
```
