---
title: "DSM Straying Model Documentation"
output: 
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{temperature-model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Chinook supplementary analysis for Chinook DSM models

### Empirical model to describe adult straying 


#### Data Source: 

This analysis is done using 2008–2011 tagging data provided by East Bay Municipal Utility District. The data can be accessed from the following links:

* [cwt_stray.csv](https://dsm-docs.s3-us-west-2.amazonaws.com/straying-data/cwt_stray_data.csv)
* [releases.csv](https://dsm-docs.s3-us-west-2.amazonaws.com/straying-data/releases.csv)
* [crxchan.csv](https://dsm-docs.s3-us-west-2.amazonaws.com/straying-data/crxchan.csv)


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(readr)
library(tidyverse)
library(DT)
```

```{r}
library(fallRunDSM)
cwt_stray_data <- read_csv(system.file("extdata", "straying-data", "cwt_stray.csv", 
                       package = "fallRunDSM", mustWork = TRUE))

release_data <- read_csv(system.file("extdata", "straying-data", "releases.csv", 
                       package = "fallRunDSM", mustWork = TRUE))

cross_channel_data <- read_csv(system.file("extdata", "straying-data", "crxchan.csv", 
                                           package = "fallRunDSM", mustWork = TRUE)) %>% 
  rename("run_year" = Run.year,
         "count" = Number,
         "prop_stray" = prop.stray,
         "pulse" = Pulsed,
         "closure_days" = Closure.days,
         "prop_sacramento" = Prp.sac,
         "prop_san_joaquin" = Prp.sj)
```

The datatable below shows the stray data that has been prepped for analysis. (`no_stray` corresponds to the number of straying fish, `prop_stray` corresponds to the proportion of straying fish, `wild` describes if fish is wild(1) or hatchery(0)
```{r}
release_data <- release_data[-2,]
combined_stray_data <- merge(cwt_stray_data, release_data, all = T)

fall <- combined_stray_data %>% 
  filter(run == "Fall") %>%
  mutate(no_tagged = ifelse(is.na(`no tagged`), 
                            no.woth.CTW, `no tagged`),
         prop_stray = ifelse(is.na(prop.stray), 0, prop.stray),
         no_stray = round(no_tagged * prop_stray),
         brood_year = as.factor(Brood.yr)) %>%
  select(no_tagged, prop_stray, no_stray, 
         brood_year, "wild" = Wild, 
         "delta" = Delta, "bay" = Bay)

datatable(fall,
          options = list(
                    searching = FALSE,
                    pageLength = 5,
                    lengthMenu = c(5, 10, 15, 20)
))
```

#### Fitting a Generalized Linear Model 

A generalized linear model is fit using the datatable above to predict the stray rate. 
** TODO:This is what they did but the results don't match up with the coefficents given in the straying submodel...there must be another straying analysis somewhere

```{r}
glm_stray_rate <- glm(cbind(no_stray, no_tagged) ~ brood_year + delta + bay + wild, 
                      family = quasibinomial, 
                      data = fall)

summary(glm_stray_rate)
```

#### Modeling Cross Channel Straying

A generalized linear model is fit using the `cross_channel_data` to predict the cross channel stray rate. 

```{r}
names(cross_channel_data)

cross_channel_data$no_stray <- round(cross_channel_data$count * cross_channel_data$prop_stray)

glm_cross_channel <- glm(cbind(no_stray, count) ~ closure_days + prop_san_joaquin, 
                         family = binomial, 
                         data = cross_channel_data)

summary(glm_cross_channel)
```

