---
title: "DSM Logistic Temperature Model Documenation"
output: 
  html_document:
    theme: flatly
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{temperature-model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.width=15, fig.height=8)
library(readr)
library(tidyverse)
library(boot) # Used for logit/inv.logit calculation
library(DT)
library(gridExtra)
```

# Chinook supplementary analysis for Chinook DSM models

## Objective 

Within the `get_rearing_survival()` DSM submodel, `aveT20`, `aveT20D`, `maxT25`, and `maxT25D` are defined as
inverse logit models whose inputs are monthly stream temperatures. The purpose of these models is to provide a sub monthly metric derived from monthly data, and these metrics are needed to calculate survival probabilities. The script below outlines the parameterization process of these models. 

## Data Sources

Central Valley River Temps: 
This analysis is done using data from the [Full temperature data.csv](https://dsm-docs.s3-us-west-2.amazonaws.com/temperature_modeling/Full+temperature+data.csv). 
The Central Valley River Temps include the following stations from USGS: 

```{r}
d <- structure(list(agency_cd = c("USGS", "USGS", "USGS", "USGS", 
"USGS", "USGS", "USGS", "USGS", "USGS", "USGS", "USGS", "USGS", 
"USGS", "USGS", "USGS", "USGS"), site_no = c("11273400", "11274550", 
"11289650", "11290000", "11303000", "11303500", "11335000", "11370500", 
"11376000", "11383500", "11389500", "11390000", "11421000", "11425500", 
"11446500", "11447650"), station_nm = c("SAN JOAQUIN R AB MERCED R NR NEWMAN CA", 
"SAN JOAQUIN R NR CROWS LANDING CA", "TUOLUMNE R BL LAGRANGE DAM NR LAGRANGE CA", 
"TUOLUMNE R A MODESTO CA", "STANISLAUS R A RIPON CA", "SAN JOAQUIN R NR VERNALIS CA", 
"COSUMNES R A MICHIGAN BAR CA", "SACRAMENTO R A KESWICK CA", 
"COTTONWOOD C NR COTTONWOOD CA", "DEER C NR VINA CA", "SACRAMENTO R A COLUSA CA", 
"BUTTE C NR CHICO CA", "YUBA R NR MARYSVILLE CA", "SACRAMENTO R A VERONA CA", 
"AMERICAN R A FAIR OAKS CA", "SACRAMENTO R A FREEPORT CA")), class = "data.frame", row.names = c(NA, 
-16L))

datatable(rename(d[2:3], `Site Number` = "site_no", `Station Name` = "station_nm"), options = list(dom = 't', pageLength = 20))
```

And the following stations from CDEC: 


```{r}
d <- structure(list(station_id = c("bic", "igo", "dvd", "grl", "crs", 
                                   "mst", "mch", "mlm", "ccr", "yrs", "bbq"), name = c("big chico creek near chico", 
                                                                                       "clear creek near igo", "deer creek below stanford vina dam", 
                                                                                       "feather river near gridley", "merced river at cressy", "merced river near stevinson", 
                                                                                       "mill creek below hwy 99", "mill creek nr los molinos", "sacramento river above clear ck", 
                                                                                       "yuba river near smartville", "stoney ck blw black butte dam"
                                   )), row.names = c(NA, -11L), class = c("tbl_df", "tbl", "data.frame"
                                   ))

datatable(rename(d, `Station ID` = station_id,  Name = name), options = list(dom = 't', pageLength = 20))
```

Delta Temps:
This analysis is done using the following data:

* [DUTCH SLOUGH.csv (USGS: 11313433 )](https://dsm-docs.s3-us-west-2.amazonaws.com/temperature_modeling/DUTCH+SLOUGH.csv)
* [GEORGIANA SLOUGH.csv (USGS: 11447903 )](https://dsm-docs.s3-us-west-2.amazonaws.com/temperature_modeling/GEORGIANA+SLOUGH.csv)
* [JERSEY POINT.csv (USGS: 11337190)](https://dsm-docs.s3-us-west-2.amazonaws.com/temperature_modeling/JERSEY+POINT.csv)
* [LPotatoSLough.csv (USGS: 11336790 )](https://dsm-docs.s3-us-west-2.amazonaws.com/temperature_modeling/LPotatoSlough.csv)

The Delta temperature datasets can be obtained using USGS `dataRetrieval` R package using the corresponding
site number listed above. For example to download the data for Jersey Point,

```r
jersey_point <- dataRetrieval::readNWISuv(siteNumbers = "11337190", parameterCd = "00010", startDate = "2009-12-11")
```


## Temperature Data 

The table below shows the Central Valley and delta temperatures grouped by month and year. It is summarized to give monthly mean temperature, the proportion of days over 20°C, and the count of days per month over 25°C. 

```{r}
full_temp_data <- read_csv(system.file("extdata", "temperature-data", "Full temperature data.csv", 
                           package = "fallRunDSM", mustWork = TRUE)) %>% 
  na.omit() %>%
  mutate(gt20 = ifelse(medC > 20, 1, 0),
         gt25 = ifelse(maxC > 25, 1, 0))

grouped_temp_measures <- full_temp_data %>%
  group_by(year, month) %>%
  summarise(prop_gt_20 = round(mean(gt20), 2),
            count_measures_gt_25 = sum(gt25),
            month_mean_temp = round(mean(medC), 2)) %>% 
  ungroup()

jersey_point <- read_csv(system.file("extdata", "temperature-data", "JERSEY POINT.csv", 
                           package = "fallRunDSM", mustWork = TRUE))
dutch_slough <- read_csv(system.file("extdata", "temperature-data", "DUTCH SLOUGH.csv", 
                           package = "fallRunDSM", mustWork = TRUE))
georgiana_slough <- read_csv(system.file("extdata", "temperature-data", "GEORGIANA SLOUGH.csv", 
                           package = "fallRunDSM", mustWork = TRUE))
potato_slough <-read_csv(system.file("extdata", "temperature-data", "LPotatoSlough.csv", 
                           package = "fallRunDSM", mustWork = TRUE))

delta_full_temp_data <- rbind(jersey_point, dutch_slough, georgiana_slough, potato_slough) %>%
  mutate(gt20 = ifelse(TempC > 20, 1, 0),
         gt25 = ifelse(TempC > 25, 1, 0))

grouped_temp_measures_delta <- delta_full_temp_data %>%
  group_by(year, month) %>%
  summarise(delta_prop_gt_20 = round(mean(gt20), 2),
            delta_count_gt_25 = sum(gt25),
            delta_month_mean_temp = round(mean(TempC), 2)) %>% 
  ungroup()

combined_temp_measures <- grouped_temp_measures_delta %>% 
  full_join(grouped_temp_measures, by = c("month", "year"))

datatable(combined_temp_measures %>% rename("Year" = year, 
                                           "Month" = month,
                                           "CV Number of Measures Exceeding 25°C" = count_measures_gt_25,
                                           "CV Proportion Exceeding 20°C" = prop_gt_20,
                                           "CV Monthly Mean Temp" = month_mean_temp,
                                           "Delta Number of Measures Exceeding 25°C" = delta_count_gt_25,
                                           "Delta Proportion Exceeding 20°C" = delta_prop_gt_20,
                                           "Delta Monthly Mean Temp" = delta_month_mean_temp),
          options = list(
                    searching = FALSE,
                    pageLength = 20,
                    lengthMenu = c(5, 10, 15, 20), 
                    dom = "tp"
))
```

## Logistic Regression models 

This data was prepped to model the proportion of days in a month with temperature over 20 °C. If the temperature goes above 25 °C, logistic regression was used to model this phenomenon. 

To allow for logistic regression model use, we made the following edits: 

* We adjusted zeros to .001 of and 1's to .999. 

* We appended an additional column to our data containing the odds ratio by taking `log(Proportion Exceeding 20/(1 - Proportion Exceeding 20))`. 

* We edited our `Number of Measures Exceeding 25 degrees` to create a binary column describing when the temperature exceeds 25 in a month (1) and when the temperature does not exceed 25 in a month (0).

```{r}
grouped_temp_measures_for_logistic <- grouped_temp_measures %>%
  mutate(prop_gt_20 = case_when(prop_gt_20 == 0 ~ 0.001, 
                                 prop_gt_20 == 1 ~ 0.999,
                                 prop_gt_20 != 0 & prop_gt_20 != 1 ~ prop_gt_20),
         odds_ratio = log(prop_gt_20 / (1 - prop_gt_20)))

#summary(grouped_temp_measures_for_logistic)
lm_summary <- summary(lm(odds_ratio ~ month_mean_temp, data = grouped_temp_measures_for_logistic))

month_means <- c(5:25)
prediction <- boot::inv.logit(lm_summary$coefficients[1, 1] + lm_summary$coefficients[2, 1] * month_means)
monthly_predictions <- data.frame(month_means, prediction)

# Repeat for delta 
delta_grouped_temp_measures_for_logit <- grouped_temp_measures_delta %>%
  ungroup() %>% 
  mutate(prop_gt_20 = case_when(delta_prop_gt_20 == 0 ~ 0.001, 
                                 delta_prop_gt_20 == 1 ~ 0.999,
                                 TRUE ~ delta_prop_gt_20),
         odds_ratio = log(prop_gt_20 / (1 - prop_gt_20)))

# summary(grouped_temp_measures_for_logit)
delta_lm_summary <- summary(lm(odds_ratio ~ delta_month_mean_temp, data = delta_grouped_temp_measures_for_logit))

month_means <- c(5:25)
delta_prediction <- boot::inv.logit(delta_lm_summary$coefficients[1, 1] + delta_lm_summary$coefficients[2, 1] * month_means)
delta_monthly_predictions <- data.frame(month_means, delta_prediction)
```

### Results: proportion of days where temperature exceedes 20°C in a month

We parameterized the model for the proportion of days where the temperature exceeds 20°C in a month by fitting a linear
regression model between the log odds of the proportion of days in a month greater than 20°C, and the monthly mean 
temperature. The logit values resulting from this are converted back into proportions using the `boot::inv.logit`.
The resulting models are, 

**Delta**

$$ prop = inv.logit(-18.18765 + 0.9560903 \times temperature)$$

**Central Valley**

$$ prop = inv.logit(-14.36524 + 0.7178891 \times temperature)$$

The scatter plots below show the mean monthly temperature and the proportion of measures greater than 20°C for the Delta and Central Valley Rivers. The red line is our logistic model prediction where the monthly temperature
ranges between 5°C and 25°C.

```{r}
cv_plot <- monthly_predictions %>%
  ggplot() +
  geom_line(aes(month_means, prediction), color = "Red") +
  geom_point(data = grouped_temp_measures_for_logistic, aes(x = month_mean_temp, y = prop_gt_20)) +
  xlab("Monthly mean temperature") +
  ylab("Proportion days temp exceeds 20°C in a month") + 
  ggtitle("Central Valley Temperature Model") +
  theme(text = element_text(size = 18))

delta_plot <- delta_monthly_predictions %>%
  ggplot() +
  geom_line(aes(month_means, prediction), color = "Red") +
  geom_point(data = delta_grouped_temp_measures_for_logit, aes(x = delta_month_mean_temp, y = delta_prop_gt_20)) +
  xlab("Monthly median temperature Delta") +
  ylab("Proportion days temp exceeds 20°C in a month") + 
  ggtitle("Delta Temperature Model") +
  theme(text = element_text(size = 18))


grid.arrange(cv_plot, delta_plot, ncol = 2)
```



```{r}
grouped_temp_measures_for_logistic <- grouped_temp_measures %>%
  mutate(count_measures_gt_25 = ifelse(count_measures_gt_25 > 1, 1, count_measures_gt_25))

glm_summary <- summary(glm(count_measures_gt_25 ~ month_mean_temp, data=grouped_temp_measures_for_logistic, family='binomial'))

month_means <- c(5:25)
prediction <- boot::inv.logit(glm_summary$coefficients[1, 1] + glm_summary$coefficients[2, 1] * month_means)
monthly_predictions <- data.frame(month_means, prediction)

# Delta model
delta_grouped_temp_measures_for_logit <- grouped_temp_measures_delta %>%
  mutate(delta_count_gt_25 = ifelse(delta_count_gt_25 > 1, 1, delta_count_gt_25))

delta_glm_summary <- summary(glm(delta_count_gt_25 ~ delta_month_mean_temp, data = delta_grouped_temp_measures_for_logit, family='binomial'))

month_means <- c(5:25) # ? where does this come from?
delta_prediction <- boot::inv.logit(glm_summary$coefficients[1, 1] + glm_summary$coefficients[2, 1] * month_means)
delta_monthly_predictions <- data.frame(month_means, delta_prediction)
```



### Results: temperatures exceeding 25°C during a month

The model for predicting when the temperature exceeds 25°C during a month is fit using a logistic regression where we regress the binary variable representing an occurrence of temperature greater than 25°C onto corresponding mean monthly temperature.

The resulting models are:

**Central Valley**

$$P_{> 25} = inv.logit(-18.66548 + 1.147803 \times temperature)$$

**Delta**

$$P_{> 25} = inv.logit(-155.563 + 6.910016 \times temperature)$$

The scatter plots below show if temp exceeds 25°C for a given mean monthly temperature in the Delta and Central Valley Rivers. The red line is our logistic model prediction based on the numeric input defined above (`boot::inv.logit(Numeric Input)`):
```{r}
cv_plot <- monthly_predictions %>%
  ggplot() +
  geom_line(aes(month_means, prediction), color = "Red") +
  geom_point(data = grouped_temp_measures_for_logistic, aes(x = month_mean_temp, y = count_measures_gt_25)) +
  xlab("Monthly mean temperature") +
  ylab("Temp goes above 25°C during month") +
  ggtitle("Central Valley Temperature Model") +
  theme(text = element_text(size = 18))

delta_plot <- delta_monthly_predictions %>%
  ggplot() +
  geom_line(aes(month_means, prediction), color = "Red") +
  geom_point(data = delta_grouped_temp_measures_for_logit, aes(x = delta_month_mean_temp, y = delta_count_gt_25)) +
  xlab("Monthly median temperature") +
  ylab("Temp goes above 25°C during month") +
  ggtitle("Delta Temperature Model") +
  theme(text = element_text(size = 18))

grid.arrange(cv_plot, delta_plot, ncol = 2) 

```





