---
title: "Juvenile Survival Submodels"
author: "Emanuel Rodriguez"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(lubridate)
library(dataRetrieval)
library(gt)
```

#### Temperature Parameters (Max temp thresh and Average temp thresh)

The following inputs are used in rearing survival, lines 145-148 in `R/survival.R`. The goal of this analysis is to produce a proportion of month that temperatures exceed a threshold (20C, 25C) based on a monthly value.

```{r, eval=FALSE}
aveT20 <- rbinom(31, 1, boot::inv.logit(-14.32252 + 0.72102 * avg_temp[ , month , year]))
maxT25 <- rbinom(31, 1, boot::inv.logit(-23.1766 + 1.4566 * avg_temp[ , month, year]))
aveT20D <- rbinom(2, 1, boot::inv.logit(-18.30017 + 0.96991 * avg_temp_delta[month, year, ]))
maxT25D <- rbinom(2, 1, boot::inv.logit(-157.537 + 6.998 * avg_temp_delta[month, year, ]))
```

##### Delta Survival

```{r, echo=FALSE}
files_to_read <- list.files("data-raw/dsm-parameter-docs/scripts-for-calculating-temp-metrics-2019_2020/", 
                            pattern = ".csv", 
                            full.names = TRUE)[-2]

# read in data and create variables for proportions 
temp_data <- map_df(files_to_read, ~read_csv(., col_types = cols(
  agency_cd = col_character(),
  site_no = col_character(),
  datetime = col_datetime(format = "%m/%d/%Y %H:%M"),
  tz_cd = col_character(),
  TempC = col_double(),
  year = col_double(),
  month = col_double(),
  day = col_double()
))) %>% 
  mutate(
    gt_20 = ifelse(TempC >= 20, 1, 0), 
    gt_25 = ifelse(TempC >= 25, 1, 0), 
    gt_25_ever = ifelse(TempC >= 25, 1, 0)
  )

```

To calcaulate the temperature proportion parameters for the delta, the following USGS Stations were used:

```{r, echo=FALSE}
sites <- temp_data %>% distinct(site_no) %>% 
  pull() %>% 
  as.character() %>% 
  map_df(~dataRetrieval::readNWISsite(.) %>% select(site_no, station_nm)) %>% 
  select(`Site Number` = site_no, 
         `Station Name` = station_nm)

gt(sites)
```

The figure below shows what these temperatures look like in 2017.

```{r}
temp_data %>%
  left_join(sites, by = c("site_no" = "Site Number")) %>% 
  filter(year(datetime) == 2017) %>% 
  ggplot(aes(datetime, TempC, color = `Station Name`)) + geom_line() + 
  theme(legend.position = "top") + 
  labs(x = "DateTime", y = "Temperature (C)", color = NULL)
```

```{r}
temp_data %>% 
  left_join(sites, by = c("site_no" = "Site Number")) %>%
  group_by(`Station Name`) %>% 
  summarise(
    `Total Observations` = format(n()), 
    `Start Date` = format(min(datetime), "%m/%d/%Y"), 
    `End Date` = format(max(datetime), "%m/%d/%Y")
  ) %>% 
  gt()
```

A time series is created by finding the proportion of times in a month when temperature exceeded either 20C or 25C at any one of the stations.

```{r}
# aggregate the data year and months
annual_agg <- 
  temp_data %>% 
  group_by(year, month) %>% 
  mutate(
    total_over_25 = sum(gt_25_ever), 
    prop_over_25 = mean(gt_25), 
    prop_over_20 = mean(gt_20), 
    monthly_mean = mean(TempC)
  ) %>% ungroup() %>% 
  select(month, year, total_over_25, prop_over_25, prop_over_20, monthly_mean) %>% 
  distinct(month, year, .keep_all = TRUE) %>% 
  arrange(year, month) %>% 
  mutate(
    prop_over_20 = ifelse(prop_over_20 == 0, 0.001, ifelse(prop_over_20 == 1, .999, prop_over_20)),
    prop_over_25 = ifelse(prop_over_25 == 0, 0.001, ifelse(prop_over_25 == 1, .999, prop_over_25)),
    total_over_25 = ifelse(total_over_25 > 1, 1, total_over_25),
    prop_eta_20 = log(prop_over_20/(1-prop_over_20)), # manual logit link
    prop_eta_25 = log(prop_over_25/(1-prop_over_25)), 
    date = as_date(paste0(year, "-", month, "-01"))
  )

annual_agg %>% 
  ggplot(aes(year, prop_over_25)) + geom_col()
```

```{r}
# get model parameters for monthly_mean = b0 + b1*prop
prop_over_20_model <- lm(prop_eta_20 ~ monthly_mean, data = annual_agg)
prop_over_25_model <- lm(prop_eta_25 ~ monthly_mean, data = annual_agg)
over_25_model <- glm(total_over_25 ~ monthly_mean, data = annual_agg, family = "binomial")

annual_agg %>% 
  ggplot(aes(monthly_mean, prop_over_20)) + geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial"))

annual_agg %>% 
  ggplot(aes(monthly_mean, total_over_25)) + geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial"))

```
